Comparing Power Optimizations in the Reorder Buffer

### 2011-06-08 paper/paper.tex

Upon execution completion, the output value is stored in the ROB with its
corresponding entry and that entry is marked as valid or completed with a
1 bit flag.  An entry is commited from the ROB to the architectrual register
file (ARF) when that entry is the oldest (lowest PC) and its result has been
marked as valid.

@@

%TODO add in bib
\begin{thebibliography}{------}
\bibitem{graphite}
MIT Graphite Simulator
https://github.com/mit-carbon/Graphite/wiki
\bibitem{kucuk}
Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Low-complexity reorder buffer architecture. In Proceedings
of the 16th international conference on Supercomputing (ICS '02). ACM, New York, NY, USA, 57-66. DOI=10.
1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
\bibitem{kucuk2}
Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Energy-Efficient Design of the Reorder Buffer. In Proceedings of the International Workshop on Power and Timing Modeling, Optimization, and Simulation, September 2002.
\bibitem{kucuk3}
Gorhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2001. Reducing Power Requirements of Instruction Scheduling Through Dynamic Allocation of Multiple Datapath Resources. In Proceedings of the International Symposium on Microarchitecture, December 2001.
\bibitem{rupnow}
Kyle Rupnow, Arun Rodrigues, Keith Underwood, and Katherine Compton. 2006. Scientific applications vs. SPEC-FP: a comparison of program behavior. In Proceedings of the 20th annual international conference on Supercomputing (ICS '06). ACM, New York, NY, USA, 66-74. DOI=10.1145/1183401.1183413 http://doi.acm.org/10.1145/1183401.1183413
\end{thebibliography}
\end{document}

>>>>

Upon execution completion, the output value is stored in the ROB with its 
corresponding entry and that entry is marked as valid or completed with a 
1 bit flag.  An entry is committed from the ROB to the architectural register
file (ARF) when that entry is the oldest (lowest PC) and its result has been 
marked as valid. 
\subsection{Circular ROB}
The circular ROB is one of the most basic implementations of the ROB.  Simply using
a circular buffer as the structure to hold each entry.  Since the ROB is basically
just memory, there are two special slots for to keep track of the head and tail
pointer to create the circular buffer.  As entries are written into the ROB, the
entry is written at the tail pointer and then the tail pointer is incremented.
Likewise when an entries is written to the ARF, the head pointer is incremented.
The circular buffer method is widely used since it inherently keeps track of the 
oldest instruction (which is at the head pointer), as well as which entry slots
are used or empty.   However it may not be very power efficient if there is a large
number of empty slots at any given point of the execution.    
\subsection{Dynamic ROB}

\subsection{Distributed ROB}

\subsection{Latch ROB}
In addition to large power requirement to maintain the entries of the ROB, another 
large power consumer is the read ports of the ROB.  The read port of the ROB is used
to forward the results of an entry to the dispatch unit (DU) to use as an operand.
All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
in an N way super scalar processor there are 2N ports (to accommodate FP
 instructions) which add unnecessary complexity and power consumption for a small 
percentage of instructions.  It has been proposed by eliminating the read ports 
in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
without read ports, a small window appears where the data cannot be accessed,
starting from when the result is placed in to the ROB until the ROB commits that 
entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
we introduce retention latches.   Retention latches are small memory elements and 
we use them to cache the most recent results that have been written into the ROB.
This way instructions are still able to access the most recently produced results
and not stall.  However if the operand needed is not in the ROB, the stream will 
stall until a commit is made to the ARF, and that result value is forwarded to 
the dispatch unit immediately.

@@

%TODO add in bib
\begin{thebibliography}{------}
\bibitem{graphite}
MIT Graphite Simulator
https://github.com/mit-carbon/Graphite/wiki
\bibitem{kucuk}
Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Low-complexity reorder buffer architecture. In Proceedings
of the 16th international conference on Supercomputing (ICS '02). ACM, New York, NY, USA, 57-66. DOI=10.
1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
\bibitem{kucuk2}
Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Energy-Efficient Design of the Reorder Buffer. In Proceedings of the International Workshop on Power and Timing Modeling, Optimization, and Simulation, September 2002.
\bibitem{kucuk3}
Gorhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2001. Reducing Power Requirements of Instruction Scheduling Through Dynamic Allocation of Multiple Datapath Resources. In Proceedings of the International Symposium on Microarchitecture, December 2001.
\bibitem{rupnow}
Kyle Rupnow, Arun Rodrigues, Keith Underwood, and Katherine Compton. 2006. Scientific applications vs. SPEC-FP: a comparison of program behavior. In Proceedings of the 20th annual international conference on Supercomputing (ICS '06). ACM, New York, NY, USA, 66-74. DOI=10.1145/1183401.1183413 http://doi.acm.org/10.1145/1183401.1183413

\end{thebibliography}
\end{document}
